An Adolescent Mini-Neptune in the Kepler Field

### 2021-08-14 paper/ms.tex
update RM

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Combining
  kinematic information from Gaia, stellar rotation periods from TESS,
  and Li abundances from ground-based spectra, we show that
  Kepler\,1627 is a member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr
  cluster.  To our knowledge, Kepler\,1627Ab is the youngest planet
  with a precise age yet found by the main Kepler mission.  From both
  unresolved starspot-crossing events in the Kepler photometry and
  from a ground-based Keck/HIRES transit we find that the planet is
  likely on an orbit nearly aligned with the stellar spin axis.  We
  also find from the short-cadence Kepler photometry that although the
  flare arrival time distribution is probably Poissonian, there may be
  an excess of flares separated by the orbital and synodic periods,
  and linear combinations thereof.  The $\delta$~Lyr cluster is one of
  many stellar groups whose properties have been clarified using the
  Gaia data; many of the other new clusters also overlap with
  confirmed and candidate exoplanets.  Further studies that combine
  Gaia kinematics and TESS rotation periods are poised to expand the
  census of age-dated planets, and our understanding of planetary
  evolution.
\end{abstract}

@@

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f16.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
    {\it Top:} 
    The time of each Kepler transit was measured, along with the local slope
    of the light curve.  The two quantities appear anti-correlated,
    which is most easily explained by starspot crossings during the
    first (second) half of transit inducing a positive (negative) TTV,
    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
    units along the abscissa are most easily understood by considering
    that the flux from the star changes by $\sim$60\,ppt per half
    rotation period ($\sim$1.3\,days).
        {\it Bottom:} Keck/HIRES transit of 7 Aug 2021 UT,
    observed simultaneously with MUSCAT3.  A quadratic baseline trend was allowed
    in the fit.
    The resulting best-fit sky-projected obliquity is listed in red.
    \label{fig:obliquity}
	}
\end{figure}

@@

{\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
Three additional new lines of evidence confirm the planetary
interpretation.  First, the $\approx$100 days of 1-minute
cadenced Kepler observations reduce the cadence-induced smearing,
enabling a more precise measurement of the transit impact parameter
\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
(3-$\sigma$) from the 1-minute cadenced data, compared to $<0.XX$
(3-$\sigma$) from the 30-minute data.

@@

The final line of evidence confirming \pn\ comes from a spectroscopic
transit we measured from Keck/HIRES during the night of 7 Aug 2021
(UT), using the iodine cell for the wavelength calibration.  We
scheduled the transit using the ephemeris of
\citet{holczer_transit_2016}, which we verified using the available
TESS data.  We also acquired simultaneous multi-band photometry using
MUSCAT3, to help interpret the radial velocities \citep{Narita_2020}.
The lower panels in Figure~\ref{fig:obliquity} show the results.  The
photometry shows the expected starspot trend, along with the transit
and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
The spectroscopic data show significant scatter, though a significant
increase in radial velocity is seen shortly after the expected
ingress.  We fitted the RVs using the \citet{hirano_analytic_2010} and
\citet{hirano_2011} models for the Rossiter-McLaughlin effect, and
considered the cases of a linear and quadratic trend.
% TODO finish this paragraph
We found $\chi^2_{\rm red} = X.XX$ for the best-fitting linear case, 
and
$\chi^2_{\rm red} = X.XX$ for the best-fitting quadratic, with a
$\Delta {\rm BIC} = Y.YY$ between the two, providing {\bf (very)
strong} evidence to prefer the latter.

>>>>

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Combining
  kinematic information from Gaia, stellar rotation periods from TESS,
  and Li abundances from ground-based spectra, we show that
  Kepler\,1627 is a member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr
  cluster.  To our knowledge, Kepler\,1627Ab is the youngest planet
  with a precise age yet found by the main Kepler mission.  From both
  unresolved starspot-crossing events in the Kepler photometry and
  from a Keck/HIRES transit we find that the planet is
  likely on an orbit nearly aligned with the stellar spin axis.  We
  also find from the short-cadence Kepler photometry that although the
  flare arrival time distribution is probably Poissonian, there may be
  an excess of flares separated by the orbital and synodic periods,
  and linear combinations thereof.  The $\delta$~Lyr cluster is one of
  many stellar groups whose properties have been clarified using the
  Gaia data; many of the other new clusters also overlap with
  confirmed and candidate exoplanets.  Further studies that combine
  Gaia kinematics and TESS rotation periods are poised to expand the
  census of age-dated planets, and our understanding of planetary
  evolution.
\end{abstract}

@@

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f16.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
    {\it Top:} 
    The time of each Kepler transit was measured, along with the local slope
    of the light curve.  The two quantities appear anti-correlated,
    which is most easily explained by starspot crossings during the
    first (second) half of transit inducing a positive (negative) TTV,
    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
    units along the abscissa are most easily understood by considering
    that the flux from the star changes by $\sim$60\,ppt per half
    rotation period ($\sim$1.3\,days).
    {\it Bottom:} Keck/HIRES transit of 7 Aug 2021 UT, observed
    simultaneously with MUSCAT3.  A quadratic baseline trend was
    allowed in the fit.  The best-fit sky-projected obliquity in this
    model is listed in red; the detection however has marginal
    statistical significance compared to a ``trend-only'' model.
    \label{fig:obliquity}
	}
\end{figure}

@@

{\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
Three additional new lines of evidence confirm the planetary
interpretation.  First, the $\approx$100 days of 1-minute
cadenced Kepler observations reduce the cadence-induced smearing,
enabling a more precise measurement of the transit impact parameter
\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
(3-$\sigma$) from the 1-minute cadenced data, compared to $<0.XX$
(3-$\sigma$) from the 30-minute data.
This implies an even stronger constraint on the transit shape in an
analysis of the false positive probability.

@@

The final line of evidence confirming \pn\ comes from a spectroscopic
transit we measured from Keck/HIRES during the night of 7 Aug 2021
(UT), using the iodine cell for the wavelength calibration.  We
scheduled the transit using the ephemeris of
\citet{holczer_transit_2016}, which we verified using the available
TESS data.  We also acquired simultaneous multi-band photometry using
MUSCAT3, to help interpret the radial velocities \citep{Narita_2020}.
The lower panels in Figure~\ref{fig:obliquity} show the results.  The
photometry shows the expected starspot trend, along with the transit
and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
The radial velocities show significant scatter, though a significant
increase in RV is seen shortly after the expected ingress.  The
internal uncertainties on the RVs ($\approx 15\,{\rm m\,s}^{-1}$) is
dominated by the $v\sin i$ broadening.

% from /drivers/logs/rm.log
% quad: 
% chi2red: 4.452
% BIC: 272.434
% AIC: 264.796
%
% linear:
% chi2red: 5.425
% BIC: 291.987
% AIC: 285.441
%
% trendonly:
% chi2red: 4.256
% BIC: 274.383
% AIC: 270.018
%
% retrograde:
% chi2red: 5.439
% BIC: 292.415
% AIC: 285.869
% 
% prograde:
% chi2red: 4.320
% BIC: 272.039
% AIC: 265.493

We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
models for the Rossiter-McLaughlin (RM) effect, which we parametrized
with three parameters: the sky-projected obliquity, the stellar
equatorial velocity, and the gaussian dispersion of spectral lines.
We then allowed for a linear and optionally quadratic trend, along
with a white-noise jitter term.  The quadratic model with the RM
effect is shown in Figure~\ref{fig:obliquity}.  We found $\chi^2_{\rm
red} = 5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} =
4.4$ for the best-fitting quadratic, with a $\Delta {\rm BIC} = {\rm
BIC}_{\rm lin} - {\rm BIC}_{\rm quad} = 19.6$ and $\Delta {\rm AIC} =
{\rm AIC}_{\rm lin} - {\rm AIC}_{\rm quad} = 20.6$.  While this
provides strong statistical evidence to prefer the quadratic model
over the linear one, the $\chi^2_{\rm red}$ indicate that neither fit
is particularly good.  Considering a quadratic trend alone without any
RM effect yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} =
{\rm BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ (similarly $\Delta
{\rm AIC}=5.2$); from a model-comparison standpoint, there is only
weak statistical evidence to prefer the best-fitting ``RM and
quadratic trend'' model over the best-fitting ``quadratic trend only''
model.  Given the expected semi-amplitude of the velocity anomaly
($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the actual achieved
precision, the claim of having ``detected the Doppler
transit'' is therefore marginal.  Observationally however, we
know from the MUSCAT3 data that the transit was observed, and the
increase in radial velocity observed shortly after ingress is more
consistent with a prograde or polar orbit than a retrograde one.
Fixing the orbit to be either exactly prograde or exactly retrograde
and allowing for a quadratic trend, we find $\Delta {\rm BIC} = {\rm
BIC}_{\rm retrograde} - {\rm BIC}_{\rm prograde} = 20.4$; the
retrograde case is ruled out on this basis.  More precise data will
however still be useful to make a more precise obliquity measurement.
